---
title: "CourseProject1"
author: "Daniel Traverso"
date: "11/1/2017"
output: html_document
---

## Loading and preprocessing the data

Read data from current working directory
``` {r read}
data <- read.csv("activity.csv")
data$date<- as.Date(data$date)
```

## What is mean total number of steps taken per day?

``` {r total per day}
data2 <- data[complete.cases(data),]
daily <- with(data2, tapply(steps, date, sum))
hist(daily, breaks=10, main="Total Daily Steps", xlab="Daily Steps")
mean(daily)
median(daily)
```

## What is the average daily activity pattern?

Times Series
``` {r time}
interval <- with(data2, tapply(steps, interval, mean))
plot(interval, type="l", xlab="Interval",ylab="Average Daily Steps")
```

Which interval contains maximum number of steps?
``` {r max interval}
interval[which.max(interval)]
```

## Imputing missing values

Number of missing values in original dataset:
``` {r missing}
sum(is.na(data$steps))
```

Impute missing values using means of each interval:
``` {r impute}
data3<- data
bad<- is.na(data3$steps)
avg_interval<- tapply(data3$steps, data3$interval, mean, na.rm=TRUE, simplify=TRUE)
data3$steps[bad] <- avg_interval[as.character(data3$interval[bad])]
data3<- data3[, c("date", "interval", "steps")]
daily2 <- with(data3, tapply(steps, date, sum))
hist(daily2, breaks=10, main="Total Daily Steps", xlab="Daily Steps")
mean(daily2)
median(daily2)
```
The mean remains the same, but now the median equals the mean.

## Are there differences in activity patterns between weekdays and weekends?

``` {r weekdays}
data3$type <- ifelse(weekdays(data3$date)=="Saturday"|weekdays(data3$date)=="Sunday","weekend","weekday")
data3$type <- as.factor(data3$type)
library(dplyr)
library(ggplot2)
interval2<- data3%>%group_by(interval,type)%>%summarise(avgStep = mean(steps))
plot<- ggplot(interval2, aes(x =interval,y=avgStep,color=type))+geom_line()+
labs(title = "Average Daily Steps by Weektype", x = "Interval", y = "Number of Steps")+facet_wrap(~type, ncol = 1, nrow=2)
print(plot)
```


The data shows more activity on the weekend.

